---
title: "chapter6"
author: "Johannes"
date: "2022-12-09"
output: html_document
---

# Chapter 6: Analysis of longitudinal data

```{r}
date()
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(corrplot)

```


#Loading RATS data and begining analyses 
```{r}
RATS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", header = TRUE, sep = '\t')
glimpse(RATS)

RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)

RATSL <- pivot_longer(RATS, cols = -c(ID, Group), 
                      names_to = "WD",
                      values_to = "Weight") %>% 
         mutate(Time = as.integer(substr(WD, 3,4))) %>%
         arrange(Time)

glimpse(RATSL)
```
#Rats data is a nutrional study with three different groups. The rats where followed for 9 weeks.

```{r}
ggplot(RATSL, aes(x = Time, y = Weight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(limits = c(min(RATSL$Weight), max(RATSL$Weight)))


```
#Here is we can see the 16 different rats by their groups. The weight of the rats increased over time within all groups. 

```{r}
# Standardising
RATSL <- RATSL %>%
  group_by(Group, Time) %>%
  mutate(stdWeight = (Weight - mean(Weight))/sd(Weight) ) %>%
  ungroup()

glimpse(RATSL)
```

```{r}
#Plotting again
ggplot(RATSL, aes(x = Time, y = stdWeight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "standardized weight")


```
#The standardized analys shows more variety in the weights of the rats.

```{r}


RATSS <- RATSL %>%
  group_by(Group, Time) %>%
  summarise(mean = mean(Weight), se = sd(Weight)/sqrt(n)) %>%
  ungroup()

glimpse(RATSS)

```

```{r}
ggplot(RATSS, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=2) +
  scale_shape_manual(values = c(1,2,3)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) +
  scale_y_continuous(name = "mean(weight) +/- se(weight)")


```

#Here we can see the mean profiles for the timepoints including standard error. Groups 2 and 3 wheight more and have a lot more variability throughout the experiment. 

```{r}
RATSL8S <- RATSL %>%
  filter(Time > 0) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()

glimpse(RATSL8S)

```

```{r}
ggplot(RATSL8S, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(Weight), Time")
```
#Lets filter the outliers 

```{r}
RATSL8S1 <- RATSL8S %>%
  filter((Group==1 & mean > 250)|(Group==2 & mean < 500)| (Group==3 & mean > 500))

ggplot(RATSL8S1, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(Weight), Time")


```

#Now lets do an Anova


```{r}
RATSL8S2 <- RATSL8S1 %>%
  mutate(baseline = RATS$WD)

fit <- lm(mean ~ Group, data = RATSL8S2)

anova(fit)

```
#As we can see the F-value is statistically significant, se we can conclude that there is a difference between the groups.

#BPRS

```{r}
#loading data and making ready for analyses

BPRS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/BPRS.txt", sep  =" ", header = T)

BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)

BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
BPRSL <-  BPRSL %>% mutate(week = as.integer(substr(weeks,5,5)))

glimpse(BPRSL)

```

```{r}
ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both)
```

#Here we can see the visualisation of 40 male participants who took part in experiment to see how two different treatments affect bprs. 

