---
title: "chapter3"
author: "Johannes"
date: "2022-11-21"
output: 
  html_document:
    theme: flatly
    highlight: haddock
    toc: true
    toc_depth: 2
    number_section: false
---

# Chapter 3

Libraries

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(GGally)
library(corrr)
library(sjPlot)

```

reading data

```{r}
alc <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/Helsinki-Open-Data-Science/master/datasets/alc.csv", sep = ",", header = TRUE)

str(alc)
colnames(alc)

```
The data has 370 obs. of  35 variables with a lot of different measures. Including alcohol use of students in secondary education of two Portuguese schools.

My hypothesis is that high freetime has a positive correlation with high alcohol use, I have the same hypothesis for absences. I'm guessing a negative correlation to alcohol if student wants to reach higher education level. I'm also hypothesising a negative correlation with having a romantic relationship

```{r}
g2 <- ggplot(alc, aes(x = high_use, y = absences,))

g2 + geom_boxplot() + ggtitle("High use and absences")

g1 <- ggplot(data = alc, aes(x = high_use))

g1 + geom_bar(aes(fill = absences))

tab_xtab(var.row = alc$high_use, var.col = alc$ab,
         title = "Title", show.row.prc = TRUE)

table(high_use = alc$high_use, absences = alc$absences) %>% prop.table() %>% addmargins()


```



```{r}

g2 <- ggplot(alc, aes(x = high_use, y = freetime,))

g2 + geom_boxplot() + ggtitle("High use and freetime")

g1 <- ggplot(data = alc, aes(x = high_use))

g1 + geom_bar(aes(fill = absences))

tab_xtab(var.row = alc$high_use, var.col = alc$freetime,
         title = "Title", show.row.prc = TRUE)


table(high_use = alc$high_use, freetime = alc$freetime) %>% prop.table() %>% addmargins()


```





```{r}

g2 <- ggplot(alc, aes(x = high_use, y = higher,))

g2 + geom_boxplot() + ggtitle("High use and higher")

g1 <- ggplot(data = alc, aes(x = high_use))

g1 + geom_bar(aes(fill = higher))

tab_xtab(var.row = alc$high_use, var.col = alc$higher,
         title = "Title", show.row.prc = TRUE)


table(high_use = alc$high_use, higher = alc$higher) %>% prop.table() %>% addmargins()

```

```{r}


g2 <- ggplot(alc, aes(x = high_use, y = romantic,))

g2 + geom_boxplot() + ggtitle("High use and romantic")

g1 <- ggplot(data = alc, aes(x = high_use))

g1 + geom_bar(aes(fill = romantic))

tab_xtab(var.row = alc$high_use, var.col = alc$romantic,
         title = "Title", show.row.prc = TRUE)


table(high_use = alc$high_use, higher = alc$romantic) %>% prop.table() %>% addmargins()


```

```{r}
test <- c("high_use","absences", "freetime", "higher", "romantic")

data2 <- select(alc, one_of(test))

glimpse(test)

p <- ggpairs(data2, mapping = aes(col = high_use, alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))

p
```



